---
title: "Auto-grouping and charting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to group sites}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(DivInsight)
library(dplyr)
data("Colombia")
```

To gain a more detailed insight of the diversity in a large area the `group_radius` argument can be used in the `clusterise_sites()` function to automatically group sites together by specifying a radius in metres.

A label is assigned to each site denoting which group it belongs to. This label can be seen in the `site_group` column in the cluster dataframe.

```{r}
# subset the dataframe by province name 
Colombia_Meta <- subset(Colombia, stateProvince == "Meta")

# cluster occurrence data by date and generate centred coordinates for each site
clusterised_Meta_20km_groups <- clusterise_sites(
  dataframe = Colombia_Meta,
  cluster_min_length = 30,
  
  group_radius = 20000
)

# view the information for each site/cluster
print(clusterised_Meta_20km_groups[[2]])
```

Base R commands can be used to see how many sites there are in each group and help us decide which group(s) to examine further.

```{r}
# use table() and sort() to see how many sites there are in each group
clusterised_Meta_20km_groups[[2]]$site_group %>% table %>% sort %>% print
```

Here we choose to examine sites 1, 5, 6, and 8 further by using `filter_groups_by_number()` to store each group's data then plot using `plot_sites_scatter_H()`.

```{r, fig.show='hold', fig.width=4, fig.height=3}
# store the data for each group
Colombia_Meta_1 <- filter_groups_by_number(clusterised_Meta_20km_groups, 1)
Colombia_Meta_5 <- filter_groups_by_number(clusterised_Meta_20km_groups, 5)
Colombia_Meta_6 <- filter_groups_by_number(clusterised_Meta_20km_groups, 6)
Colombia_Meta_8 <- filter_groups_by_number(clusterised_Meta_20km_groups, 8)

# plot the data in a scatter plot
plot_sites_scatter_H(Colombia_Meta_1, main = "Shannons H at Meta #1") 
plot_sites_scatter_H(Colombia_Meta_5, main = "Shannons H at Meta #5") 
plot_sites_scatter_H(Colombia_Meta_6, main = "Shannons H at Meta #6") 
plot_sites_scatter_H(Colombia_Meta_8, main = "Shannons H at Meta #8") 
```

The charts show varying patterns of diversity over available time frames but it is important to know where these changes are taking place.

`map_start()` can be used to create a new map object and `map_add()` can be used to add coordinate information. 

```{r}
# create a new map using group 1's coordinates
Colombia_Meta_map <- 
  
  map_start(
    
    clusterised_object = Colombia_Meta_1,
    site_name = "Meta #1",
    colour = "green"
  )

# add group 5's coordinates to the map object 
Colombia_Meta_map <-
  
  map_add(
    
    existing_map = Colombia_Meta_map,
    
    clusterised_object = Colombia_Meta_5,
    site_name = "Meta #5",
    colour = "purple"
  )

# add group 6's coordinates to the map object
Colombia_Meta_map <-
  
  map_add(
    
    existing_map = Colombia_Meta_map,
    
    clusterised_object = Colombia_Meta_6,
    site_name = "Meta #6",
    colour = "blue"
  )

# add group 8's coordinates to the map object
Colombia_Meta_map <-
  
  map_add(
    
    existing_map = Colombia_Meta_map,
    
    clusterised_object = Colombia_Meta_8,
    site_name = "Meta #8",
    colour = "red"
  )
```

Once the map has been created it can be viewed.

```{r, fig.show='hold', fig.width=8, fig.height=6}
# view the map
Colombia_Meta_map
```
